---
title: "Curlee Milestone_4"
format: html
editor: visual
---


# Combined Database

```{r}
#| include: false
#| warning: false
#| message: false

# Load and install packages
pacman::p_load(
  dplyr, tidyr, ggplot2, lubridate, ggthemes, cowplot, readr, rlang, purrr, rlang,
  classInt, tidycensus, sf, here, stringr, purrr, svglite, rmapshaper, readxl,
  scales, ggrepel, viridis, RColorBrewer, maps, ggfx, glue, skimr, DataExplorer,
  knitr, kableExtra, janitor, reactable, MMWRweek
)

# Source all helper functions
func_dir <- here::here("functions")
r_files <- list.files(func_dir, pattern = "\\.R$", full.names = TRUE)
purrr::walk(r_files, source)


# === Load all CSVs from _data/tbl_outputs_03 ===
data_dir <- here::here("data", "tbl_outputs_03")

csv_files <- list.files(
  path = data_dir,
  pattern = "\\.csv$",
  full.names = TRUE
)

# Read and assign each CSV to an object named after its file
purrr::walk(csv_files, function(file_path) {
  # Extract clean base name (no extension)
  obj_name <- tools::file_path_sans_ext(basename(file_path))
  # Read file
  df <- readr::read_csv(file_path, show_col_types = FALSE)
  # Assign to the global environment so all chapters can access
  assign(obj_name, df, envir = knitr::knit_global())
})


```


After reconciling all column names across the 3 datasets, standardizing age group, race/ethnicity, and infection date variables, we merge them together to generate a final, complete database and begin our descriptive explorations and analyses. The final, combined database as well as the cleaned individual datasets (California, Los Angeles, and population data) are stored and saved for future reference. 


[Download Combined Database (CSV)](../../data/cleaned_data/combined_df.csv){.btn .btn-primary target="_blank"}

```{r, eval = FALSE}

step2_ca_df <- read.csv(file = here("data/tbl_outputs_03/step2_ca_df.csv"))
step2_la_cnty_df <- read.csv(file = here("data/tbl_outputs_03/step2_la_cnty_df.csv"))
step2_pop_df <- read.csv(file = here("data/tbl_outputs_03/step2_pop_df.csv"))


combined_df <- rbind(step2_ca_df, step2_la_cnty_df) %>%
  relocate(health_officer_region,  .after = county) %>%
  relocate(pop, .after = race_long) %>%
  mutate(age_cat = factor(age_cat, levels = c("0-17", "18-49", "50-64", "65+")))

```


**Glimpse of database structure:**

```{r}
#| code-fold: show

dplyr::glimpse(combined_df)

```


```{r}
step2_pop_df <- step1_pop_df %>%
  group_by(
      county, health_officer_region, 
      sex, race_ethnicity, age_cat
  ) %>%
  summarise(pop = sum(pop), .groups = "drop")

#-aggregate population counts 
##--to 4 age category format:
step2_pop_df_recat <- step1_pop_df %>%
  mutate(
    new_age_group = 
      if_else(
        age_cat %in% 
          c("0-4", "5-11", "12-17"),
              "0-17", age_cat)
  ) %>%
  group_by(
      county, health_officer_region, sex, 
      race_ethnicity, new_age_group
  ) %>%
  summarise(pop = sum(pop), .groups = "drop") %>%
  rename("age_cat" = "new_age_group")
  

```

```{r}
race_ethnicity_map <- read.csv(file = here("data/race_ethnicity_map.csv")) %>%
  mutate(
    race_long = clean(race_long),
    race_short = clean(race_short),
    race_coded = as.character(race_coded)
  )

```


```{r}
step2_pop_df <- step2_pop_df_recat %>%
  
##-- join population database to the race/ethnicity map
  mutate(race_short = clean(race_ethnicity)) %>%
  select(-race_ethnicity) %>%
  left_join(race_ethnicity_map, by = "race_short") %>%
  relocate(race_coded, race_short, race_long, .after = sex) %>%
  
##-- calculate population totals by demographic
  group_by(county, health_officer_region) %>%
    mutate(total_cnty_pop = sum(pop)) %>% ungroup() %>%
  group_by(county, health_officer_region, race_coded, race_short, race_long) %>%
    mutate(total_race_pop = sum(pop)) %>% ungroup() %>%
  group_by(county, health_officer_region, sex) %>%
    mutate(total_sex_pop = sum(pop)) %>% ungroup() %>%
  group_by(health_officer_region) %>%
    mutate(total_HOR_pop = sum(pop)) %>% ungroup()

```



```{r}
by_sex <- combined_df %>%
  group_by(health_officer_region, sex) %>%
  summarise(
    total_pop = sum(total_cnty_pop, na.rm = TRUE),
    total_infected = sum(cumulative_infected, na.rm = TRUE),
    prop_infected = total_infected / total_pop,
    rate_100k = round((prop_infected * 1e5),1),
    .groups = "drop"
  )


tot_by_sex <- by_sex %>%
  group_by(health_officer_region) %>%
  summarise(
    sex = "BOTH",
    total_pop = sum(total_pop, na.rm = TRUE),
    total_infected = sum(total_infected, na.rm = TRUE),
    prop_infected = total_infected/total_pop,
    rate_100k = round(prop_infected * 1e5,1),
    .groups = "drop")

 

by_sex_tbl <- by_sex %>%
  filter(sex %in% c("FEMALE", "MALE")) %>% 
  bind_rows(tot_by_sex) %>% 
  arrange(health_officer_region,
          factor(sex, levels = c("FEMALE", "MALE", "BOTH")))

sex_tbl <- by_sex_tbl %>% 
  arrange(health_officer_region, sex) %>% 
  mutate(
    Region = ifelse(duplicated(health_officer_region), "", health_officer_region),
    sex = paste0(" ", sex))

sex_tbl %>% 
  select(
    Region,
    sex,
    Total_Population = total_pop,
    Total_Infected = total_infected,
    Proportion_Infecteed = prop_infected,
    Infection_Rate_100K = rate_100k) %>% 
  
kbl(
    escape = FALSE, 
    align = "c",
    col.names = 
      c("Health Officer Region", "", "Total Population", 
        "Total Infected", "Proportion Infected", 
        "Infection Rate per 100K")
) %>%
row_spec(0, 
       bold = TRUE, 
       background = "#0f172a",
       extra_css = "font-size: 16px!important;color:#ffffff;"
) %>%
  
column_spec(6, 
       bold = TRUE, 
       background = "#5ce1e6"
) %>%  
kable_styling(bootstrap_options = c("bordered"))

```


```{r}
```


```{r}
```








