# Infection Rates by Demographic


```{r}
#| include: false
#| warning: false
#| message: false

source("_common.R")

```


```{r}
#| code-fold: show

# Bring in final combined database
df <- combined_df

```




## Calculating Crude Rates

To simplify and standardize the process of calculating crude infection rates across multiple demographic and geographic groupings, we created a custom helper function: `calc_rates_rng_grp()` 

It ensures consistent calculations across datasets and minimizes the need to repeat long dplyr pipes for each demographic of interest.

Please see the [Custom Functions Appendix](apx_custom_functions.qmd#custom-fx) and the 
[Supporting Code](#supporting-code) section below on this page for details and documentation.




#### Infection Rates by Demographic and Region, Counties
##### *California Outbreak, May 2023 - Dec 2023*

```{r, eval = FALSE}

##########################
#--Prepare dataframe
##--see 04i_calc_rates_rng_grp.R function
###--in 'functions/' folder

# specs: which grouping var uses which population column
group_specs <- tibble::tibble(
  group_var = list(quo(sex), quo(age_cat), quo(race_short)),
  pop_col   = list(quo(total_sex_pop), quo(total_age_pop), quo(total_race_pop))
)

############################
##  Region-level (HOR)   ###
############################

hor_results_list <- group_specs %>%
  purrr::pmap(function(group_var, pop_col) {
    calc_rates_rng_grp(
      df        = df,
      group_var = !!group_var,
      pop_col   = !!pop_col,
      geo_level = "region"   
    ) %>%
      dplyr::mutate(geo_level = "region")
  })

hor_rates_by_demographic <- dplyr::bind_rows(hor_results_list)

############################
##      County-level      ##
############################

cnty_results_list <- group_specs %>%
  purrr::pmap(function(group_var, pop_col) {
    calc_rates_rng_grp(
      df        = df,
      group_var = !!group_var,
      pop_col   = !!pop_col,
      geo_level = "county"
    ) %>%
      dplyr::mutate(geo_level = "county")
  })

cnty_rates_by_demographic <- dplyr::bind_rows(cnty_results_list)



############################
##       Statewide        ##
############################

statewide_rates_by_demographic <- group_specs %>%
  purrr::pmap(function(group_var, pop_col) {
    calc_rates_rng_grp(
      df        = combined_df,
      group_var = !!group_var,
      pop_col   = !!pop_col,
      geo_level = "statewide"
    ) %>%
      dplyr::mutate(geo_level = "statewide")
  }) %>%
  dplyr::bind_rows()



#########################################
## - calculates total rates by geography 
###--(no stratification)
geo_rates_overall <- calc_rates_overall_hor_cnty(df)


############################################
##   all rates by dem combined in one df  ##
############################################

all_rates_by_demographic <- dplyr::bind_rows(
  hor_rates_by_demographic,
  cnty_rates_by_demographic,
  statewide_rates_by_demographic,
  geo_rates_overall
)
  
#########################################
## Exports to 'data/cleaned_data/' dir ##
#########################################

#write.csv(cnty_rates_by_demographic, here("data/cleaned_data/cnty_rates_by_demographic.csv"), row.names = FALSE)
#write.csv(hor_rates_by_demographic, here("data/cleaned_data/hor_rates_by_demographic.csv"), row.names = FALSE)
#write.csv(all_rates_by_demographic, here("data/cleaned_data/all_rates_by_demographic.csv"), row.names = FALSE)

```




```{r}
#| code-fold: show


dem_df <- all_rates_by_demographic


```




















```{r, echo = FALSE}

HOR_filt <- make_rate_table(
  by_HOR,
  health_officer_region,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = NULL,                 
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)

cnty_filt <- make_rate_table(
  by_cnty,
  county,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = 10,
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)


age_filt <- make_rate_table(
  by_age,
  age_cat,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = NULL,                 
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)


sex_filt <- make_rate_table(
  by_sex,
  sex,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = NULL,                 
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)


race_filt <- make_rate_table(
  by_race,
  race_long,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = NULL,                 
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)

```


## Severe Infection Rates by Demographic


::: {.panel-tabset}


## Health Officer Region

```{r, echo = FALSE}
HOR_filt
```



## County

**Top 10 Highest Severe Infection Rates by County**
```{r, echo = FALSE}
cnty_filt
```

## Age Category

```{r, echo = FALSE}
age_filt
```


## Sex

*initial note: males have higher infection rate; but females have higher rates of severe infections and unrecovered*

```{r, echo = FALSE}
sex_filt
```



## Race/Ethnicity

*initial note: Hispanic population has highest infection rate, severe infection rate, and unrecovered rate -- all across the board -- this community is being impacted the most.*

```{r, echo = FALSE}
race_filt
```



:::









|
|
|
|


#### Supporting Code {#supporting-code}


*Custom function to calculate infection rates by geographic and demographic group(s):*


```{r, eval = FALSE}


####################################################
##                                                ##
##  Calculate rates based on group and geography  ##
##     (health officer region or county)          ##
##                                                ##
####################################################

calc_rates_rng_grp  <- function(df,
                                group_var,      
                                pop_col,        
                                week        = 52,
                                region_col  = health_officer_region,
                                county_col  = county,
                                infected_col = cumulative_infected,
                                severe_col   = cumulative_severe,
                                geo_level    = c("region", "county", "statewide")) {
  
  group_var    <- rlang::enquo(group_var)
  pop_col      <- rlang::enquo(pop_col)
  region_col   <- rlang::enquo(region_col)
  county_col   <- rlang::enquo(county_col)
  infected_col <- rlang::enquo(infected_col)
  severe_col   <- rlang::enquo(severe_col)
  geo_level    <- match.arg(geo_level)
  
  group_var_name <- rlang::as_name(group_var)
  
  #-- get *all* possible group levels (e.g. all 7 race categories)
  all_groups <- df %>%
    dplyr::pull(!!group_var) %>%
    unique()
  
  #-- restrict to last week only (defaults to 52)
  df_week <- df %>%
    dplyr::filter(mmwr_week == week)
  
  #-- geography-specific aggregation
  out <- if (geo_level == "region") {
    
#-- health officer region-level
    df_week %>%
      dplyr::group_by(!!region_col, !!county_col, !!group_var) %>%
      dplyr::summarise(
        group_pop           = dplyr::first(!!pop_col),
        cumulative_infected = sum(!!infected_col, na.rm = TRUE),
        cumulative_severe   = sum(!!severe_col,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(!!region_col, !!group_var) %>%
      dplyr::summarise(
        group_pop           = sum(group_pop,           na.rm = TRUE),
        cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
        cumulative_severe   = sum(cumulative_severe,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(!!region_col) %>%
      tidyr::complete(
        !!group_var := all_groups,
        fill = list(
          group_pop           = 0,
          cumulative_infected = 0,
          cumulative_severe   = 0
        )
      ) %>%
      dplyr::ungroup()
    
  } else if (geo_level == "county")  {
    
#--county-level
    df_week %>%
      dplyr::group_by(!!region_col, !!county_col, !!group_var) %>%
      dplyr::summarise(
        group_pop           = sum(!!pop_col,      na.rm = TRUE),
        cumulative_infected = sum(!!infected_col, na.rm = TRUE),
        cumulative_severe   = sum(!!severe_col,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(!!region_col, !!county_col) %>%
      tidyr::complete(
        !!group_var := all_groups,
        fill = list(
          group_pop           = 0,
          cumulative_infected = 0,
          cumulative_severe   = 0
        )
      ) %>%
      dplyr::ungroup()
    
    
  } else {
    
#--statewide (no county or hor grouping) 
    
    df_week %>%
  ##-- first collapse to one row per county and group
      dplyr::group_by(!!county_col, !!group_var) %>%
      dplyr::summarise(
        group_pop_county    = dplyr::first(!!pop_col),
        cumulative_infected = sum(!!infected_col, na.rm = TRUE),
        cumulative_severe   = sum(!!severe_col,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
  ##-- then sum across counties to get statewide totals by group
      dplyr::group_by(!!group_var) %>%
      dplyr::summarise(
        group_pop           = sum(group_pop_county,    na.rm = TRUE),
        cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
        cumulative_severe   = sum(cumulative_severe,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
      tidyr::complete(
        !!group_var := all_groups,
        fill = list(
          group_pop           = 0,
          cumulative_infected = 0,
          cumulative_severe   = 0
        )
      ) %>%
  ##-- adds "statewide" for joining compatibility
      dplyr::mutate(
        !!region_col := "statewide",
        !!county_col := "statewide"
      )
  }   
    

#################  
#-- compute rates
  out %>%
    dplyr::mutate(
      inf_rate_100k = dplyr::if_else(
        group_pop > 0,
        round(1e5 * cumulative_infected / group_pop),
        0
      ),
      sev_rate_100k = dplyr::if_else(
        group_pop > 0,
        round(1e5 * cumulative_severe / group_pop),
        0
      )
    ) %>%
    dplyr::rename(group_var_cat = !!group_var) %>%
    dplyr::mutate(group_var = group_var_name) %>%
    relocate(group_var, .after = group_var_cat) %>%
    
##--These last two columns added are just a QC checkpoint:
###-- if grouping worked properly, population sums 
###---should all equal total California population (39109070)
    
    group_by(group_var) %>%
      mutate(
        total_group_var_pop = sum(group_pop),
               total_ca_pop = 39109070
        ) %>%
    ungroup()
}

```









**Code for table subsets:**


```{r, eval = FALSE}

HOR_filt <- make_rate_table(
  by_HOR,
  health_officer_region,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = NULL,                 
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)

cnty_filt <- make_rate_table(
  by_cnty,
  county,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = 10,
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)


age_filt <- make_rate_table(
  by_age,
  age_cat,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = NULL,                 
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)


sex_filt <- make_rate_table(
  by_sex,
  sex,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = NULL,                 
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)


race_filt <- make_rate_table(
  by_race,
  race_long,
  pop_col = total_pop,
  sort_by = "sev_rate_100k",
  top_n = NULL,                 
  drop_cols = "unrec_rate_100k",
  highlight_col = "sev_rate_100k"
)

```
