# Combined Demographic Dataset



```{r, eval = FALSE}
#| code-fold: show

source("_common.R")

# Bring in final combined database
df <- combined_df

```




## Calculating Crude Rates

To simplify and standardize the process of calculating crude infection rates across multiple demographic and geographic groupings, we created a custom helper function: `calc_rates_rng_grp()` 

It ensures consistent calculations across datasets and minimizes the need to repeat long dplyr pipes for each demographic of interest.

Please see the [Custom Functions Appendix](../apx_custom_functions.qmd#custom-fx) for more details and documentation of this function. 




*Custom function to calculate infection rates by geographic and demographic group(s):*


```{r, eval = FALSE}


####################################################
##                                                ##
##  Calculate rates based on group and geography  ##
##     (health officer region or county)          ##
##                                                ##
####################################################

calc_rates_rng_grp  <- function(df,
                                group_var,      
                                pop_col,        
                                week        = 52,
                                region_col  = health_officer_region,
                                county_col  = county,
                                infected_col = cumulative_infected,
                                severe_col   = cumulative_severe,
                                geo_level    = c("region", "county", "statewide")) {
  
  group_var    <- rlang::enquo(group_var)
  pop_col      <- rlang::enquo(pop_col)
  region_col   <- rlang::enquo(region_col)
  county_col   <- rlang::enquo(county_col)
  infected_col <- rlang::enquo(infected_col)
  severe_col   <- rlang::enquo(severe_col)
  geo_level    <- match.arg(geo_level)
  
  group_var_name <- rlang::as_name(group_var)
  
  #-- get *all* possible group levels (e.g. all 7 race categories)
  all_groups <- df %>%
    dplyr::pull(!!group_var) %>%
    unique()
  
  #-- restrict to last week only (defaults to 52)
  df_week <- df %>%
    dplyr::filter(mmwr_week == week)
  
  #-- geography-specific aggregation
  out <- if (geo_level == "region") {
    
#-- health officer region-level
    df_week %>%
      dplyr::group_by(!!region_col, !!county_col, !!group_var) %>%
      dplyr::summarise(
        group_pop           = dplyr::first(!!pop_col),
        cumulative_infected = sum(!!infected_col, na.rm = TRUE),
        cumulative_severe   = sum(!!severe_col,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(!!region_col, !!group_var) %>%
      dplyr::summarise(
        group_pop           = sum(group_pop,           na.rm = TRUE),
        cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
        cumulative_severe   = sum(cumulative_severe,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(!!region_col) %>%
      tidyr::complete(
        !!group_var := all_groups,
        fill = list(
          group_pop           = 0,
          cumulative_infected = 0,
          cumulative_severe   = 0
        )
      ) %>% ungroup() %>%
      mutate(county = NA_character_) %>%
      relocate(county, .after = health_officer_region)
      
    
  } else if (geo_level == "county")  {
    
#--county-level
    df_week %>%
      dplyr::group_by(!!region_col, !!county_col, !!group_var) %>%
      dplyr::summarise(
        group_pop           = first(!!pop_col),
        cumulative_infected = sum(!!infected_col, na.rm = TRUE),
        cumulative_severe   = sum(!!severe_col,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(!!region_col, !!county_col) %>%
      tidyr::complete(
        !!group_var := all_groups,
        fill = list(
          group_pop           = 0,
          cumulative_infected = 0,
          cumulative_severe   = 0
        )
      ) %>%
      dplyr::ungroup()
    
    
  } else {
    
#--statewide (no county or hor grouping) 
    
    df_week %>%
  ##-- first collapse to one row per county and group
      dplyr::group_by(!!county_col, !!group_var) %>%
      dplyr::summarise(
        group_pop_county    = dplyr::first(!!pop_col),
        cumulative_infected = sum(!!infected_col, na.rm = TRUE),
        cumulative_severe   = sum(!!severe_col,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
  ##-- then sum across counties to get statewide totals by group
      dplyr::group_by(!!group_var) %>%
      dplyr::summarise(
        group_pop           = sum(group_pop_county,    na.rm = TRUE),
        cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
        cumulative_severe   = sum(cumulative_severe,   na.rm = TRUE),
        .groups = "drop"
      ) %>%
      tidyr::complete(
        !!group_var := all_groups,
        fill = list(
          group_pop           = 0,
          cumulative_infected = 0,
          cumulative_severe   = 0
        )
      ) %>%
  ##-- adds "statewide" for joining compatibility
      dplyr::mutate(
        !!region_col := "statewide",
        !!county_col := "statewide"
      )
  }   
    

#################  
#-- compute rates
  out %>%
    dplyr::mutate(
      inf_rate_100k = dplyr::if_else(
        group_pop > 0,
        round(1e5 * cumulative_infected / group_pop),
        0
      ),
      sev_rate_100k = dplyr::if_else(
        group_pop > 0,
        round(1e5 * cumulative_severe / group_pop),
        0
      )
    ) %>%
    dplyr::rename(group_var_cat = !!group_var) %>%
    dplyr::mutate(group_var = group_var_name) %>%
    relocate(group_var, .after = group_var_cat) %>%
    
##--These last two columns added are just a QC checkpoint
###-- if grouping worked properly, population sums 
###---should all equal total California population (39109070)
    
    group_by(group_var) %>%
      mutate(
        total_group_var_pop = sum(group_pop),
               total_ca_pop = 39109070
        ) %>%
    ungroup()
}

```


|
|


*Create final, combined dataset with all rates by demographic and geography:*


```{r, eval = FALSE}


##########################
#--Prepare dataframe
##--see 04i_calc_rates_rng_grp.R function
###--in 'functions/' folder

# specs: which grouping var uses which population column
group_specs <- tibble::tibble(
  group_var = list(quo(sex), quo(age_cat), quo(race_short)),
  pop_col   = list(quo(total_sex_pop), quo(total_age_pop), quo(total_race_pop))
)

############################
##  Region-level (HOR)    ##
############################

hor_results_list <- group_specs %>%
  purrr::pmap(function(group_var, pop_col) {
    calc_rates_rng_grp(
      df        = df,
      group_var = !!group_var,
      pop_col   = !!pop_col,
      geo_level = "region"   
    ) %>%
      dplyr::mutate(geo_level = "region")
  })

hor_rates_by_demographic <- dplyr::bind_rows(hor_results_list)

############################
##      County-level      ##
############################

cnty_results_list <- group_specs %>%
  purrr::pmap(function(group_var, pop_col) {
    calc_rates_rng_grp(
      df        = df,
      group_var = !!group_var,
      pop_col   = !!pop_col,
      geo_level = "county"
    ) %>%
      dplyr::mutate(geo_level = "county")
  })

cnty_rates_by_demographic <- dplyr::bind_rows(cnty_results_list)



############################
##       Statewide        ##
############################

statewide_rates_by_demographic <- group_specs %>%
  purrr::pmap(function(group_var, pop_col) {
    calc_rates_rng_grp(
      df        = combined_df,
      group_var = !!group_var,
      pop_col   = !!pop_col,
      geo_level = "statewide"
    ) %>%
      dplyr::mutate(geo_level = "statewide")
  }) %>%
  dplyr::bind_rows()



#########################################
## - calculates total rates by geography 
###--(no stratification)
geo_rates_overall <- calc_rates_overall_hor_cnty(df)


############################################
##   all rates by dem combined in one df  ##
############################################

all_rates_by_demographic <- dplyr::bind_rows(
  hor_rates_by_demographic,
  cnty_rates_by_demographic,
  statewide_rates_by_demographic,
  geo_rates_overall
)
  
#########################################
## Exports to 'data/cleaned_data/' dir ##
#########################################

#write.csv(cnty_rates_by_demographic, here("data/cleaned_data/cnty_rates_by_demographic.csv"), row.names = FALSE)
#write.csv(hor_rates_by_demographic, here("data/cleaned_data/hor_rates_by_demographic.csv"), row.names = FALSE)
#write.csv(all_rates_by_demographic, here("data/cleaned_data/all_rates_by_demographic.csv"), row.names = FALSE)

```



|
|


## Calculating Age-Adjusted Rates


We next calculate age and race adjusted severe infection rates to make fair comparisons across counties that have different population structures. 

* First, we use the statewide distribution of race and age groups as our standard and calculate stratum specific severe infection rates within each county. 

* Then, we then weight each county’s race or age specific rate by the corresponding statewide proportion and sum these weighted rates to obtain age adjusted, race adjusted, and jointly age and race adjusted severe infection rates per 100,000. 

* Finally, we create a new combined dataset of rates by demographic, with the new columns for the adjusted infection rates. 



*Calculate age, sex,  and race-adjusted severe infection rates, append and overwrite final demographic dataset:*



```{r, eval = FALSE}

### -- had to run age x sex separately then will join
##### by age x sex 

#################################
######### statewide rates
by_sex_age_state <- combined_df %>% 
  #--only last week (52)
  filter(mmwr_week == 52) %>%
  #--one row per region–county–sex 
  group_by(mmwr_week, sex, age_cat) %>%
  summarise(
    sex_age_pop         = sum(pop),
    total_ca_pop        = first(total_ca_pop),
    cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
    cumulative_severe   = sum(cumulative_severe, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(sex) %>%
  mutate(total_sex_pop = sum(sex_age_pop)) %>% 
  ungroup() %>%
  mutate(
    age_sex_prop        = sex_age_pop/total_sex_pop,
    inf_rate_100k       = round((10^5*cumulative_infected/sex_age_pop)),
    sev_rate_100k       = round((10^5*cumulative_severe/sex_age_pop))
  )


std_age_weights_sex <- by_sex_age_state %>%
  group_by(sex, age_cat) %>%
  summarise(
    w = sex_age_pop / total_sex_pop,
    .groups = "drop"
  )


by_sex_age_cnty <- combined_df %>% 
  #--only last week (52)
  filter(mmwr_week == 52) %>%
  #--one row per region–county–sex 
  group_by(health_officer_region, county, age_cat, sex) %>%
  summarise(
    sex_age_pop         = sum(pop),
    total_cnty_pop      = first(total_cnty_pop),
    cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
    cumulative_severe   = sum(cumulative_severe, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    inf_rate       = cumulative_infected/sex_age_pop,
    sev_rate       = cumulative_severe/sex_age_pop
  ) 


cnty_sex_age_adj <- by_sex_age_cnty %>%
  left_join(std_age_weights_sex, by = c("sex", "age_cat")) %>%
  group_by(health_officer_region, county, sex) %>%
  summarise(
    adj_rate_100k = 1e5 * sum(inf_rate * w, na.rm = TRUE),
    adj_sev_100k = 1e5 * sum(sev_rate * w, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(county) %>%

  ungroup() %>%
  mutate(group_var = "sex", geo_level = "county") %>%
  rename(group_var_cat = sex)


```


```{r, eval = FALSE}


####################################################
##                                                ##
##     Calculate age, sex, and race-adjusted      ##
##          severe rates by county                ##
##                                                ##
####################################################

std_race_weights <- all_rates_by_demographic %>%
  filter(
    geo_level == "statewide",
    group_var == "race_short"
  ) %>%
  group_by(group_var_cat) %>%
  summarise(std_pop = sum(group_pop), .groups = "drop") %>%
  mutate(w = std_pop / sum(std_pop)) %>%
  ungroup()




cnty_race_rates <- all_rates_by_demographic %>%
  filter(
    geo_level == "county",
    group_var == "race_short"
  ) %>%
  group_by(county, group_var_cat) %>%
  summarise(
    infected = sum(cumulative_infected, na.rm = TRUE),
    pop    = sum(group_pop, na.rm = TRUE),
    rate   = infected / pop,
    .groups = "drop"
  )


cnty_race_adj <- cnty_race_rates %>%
  left_join(std_race_weights, by = "group_var_cat") %>%
  group_by(county) %>%
  summarise(
    adj_rate_100k = 10^5 *(sum(rate * w, na.rm = TRUE)),
    .groups       = "drop"
  )%>%
  mutate(
    group_var = "race_short",
    geo_level = "county"
  )

cnty_race_rates_sev <- all_rates_by_demographic %>%
  filter(
    geo_level == "county",
    group_var == "race_short"
  ) %>%
  group_by(county, group_var_cat) %>%
  summarise(
    severe = sum(cumulative_severe, na.rm = TRUE),
    pop    = sum(group_pop, na.rm = TRUE),
    rate   = severe / pop,
    .groups = "drop"
  )


cnty_race_adj_sev <- cnty_race_rates_sev %>%
  left_join(std_race_weights, by = "group_var_cat") %>%
  group_by(county) %>%
  summarise(
    adj_sev_100k = 10^5 *(sum(rate * w, na.rm = TRUE)),
    .groups       = "drop"
  )%>%
  mutate(
    group_var = "race_short",
    geo_level = "county"
  )

cnty_adj_race_df <- cnty_race_adj_sev %>%
  left_join(cnty_race_adj, by = c("county", "group_var", "geo_level"))

######################################################
### age-adjusted severe rates
## - uses CA pop distribution by age group as standard

std_age_weights <- all_rates_by_demographic %>%
  filter(
    geo_level == "statewide",
    group_var == "age_cat"
    ) %>%
  group_by(group_var_cat) %>%
  summarise(std_pop = sum(group_pop), .groups = "drop") %>%
  mutate(w = std_pop / sum(std_pop)) %>% 
  ungroup()


cnty_age_sev_rates <- all_rates_by_demographic %>%
  filter(
    geo_level == "county",
    group_var == "age_cat"
  ) %>%
  group_by(county, group_var, group_var_cat) %>%
  summarise(
    severe = sum(cumulative_severe, na.rm = TRUE),
    pop    = sum(group_pop, na.rm = TRUE),
    rate   = severe / pop,
    .groups = "drop"
  )


cnty_age_adj_sev <- cnty_age_sev_rates %>%
  left_join(std_age_weights, by = "group_var_cat") %>%
  group_by(county) %>%
  summarise(
    adj_sev_100k = 10^5 *(sum(rate * w, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    group_var = "age_cat",
    geo_level = "county"
  )


cnty_age_rates <- all_rates_by_demographic %>%
  filter(
    geo_level == "county",
    group_var == "age_cat"
  ) %>%
  group_by(county, group_var, group_var_cat) %>%
  summarise(
    infected = sum(cumulative_infected, na.rm = TRUE),
    pop      = sum(group_pop, na.rm = TRUE),
    rate     = infected / pop,
    .groups  = "drop"
  )

cnty_age_adj <- cnty_age_rates %>%
  left_join(std_age_weights, by = "group_var_cat") %>%
  group_by(county) %>%
  summarise(
    adj_rate_100k = 10^5 *(sum(rate * w, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    group_var = "age_cat",
    geo_level = "county"
  )


cnty_adj_case_df <- cnty_age_adj_sev %>%
  left_join(cnty_age_adj, by = c("county", "group_var", "geo_level"))


##--joint adjustment: age cat + race

ca_age_race_rates <- df %>%
  filter(mmwr_week == 52) %>%
  group_by(age_cat, race_short) %>%
  summarise(
    pop = sum(pop, na.rm =  TRUE), 
    severe = sum(cumulative_severe, na.rm= TRUE),
    rate   = severe / pop,
    .groups = "drop"
    ) 

std_age_race_weights <- ca_age_race_rates %>%
  group_by(age_cat, race_short) %>%
  summarise(std_pop = sum(pop), .groups = "drop") %>%
  mutate(w = std_pop / sum(std_pop)) %>%
  ungroup()


cnty_age_race_rates <- df %>%
  filter(mmwr_week == 52) %>%
  group_by(county, age_cat, race_short) %>%
  summarise(
    pop = sum(pop, na.rm =  TRUE), 
    severe = sum(cumulative_severe, na.rm= TRUE),
    rate   = severe / pop,
    .groups = "drop"
    ) 


cnty_joint_adj <- cnty_age_race_rates %>%
  left_join(std_age_race_weights, 
            by = c("age_cat", "race_short")) %>%
  group_by(county) %>%
  summarise(
    adj_sev_joint_rate = 10^5 *(sum(rate * w, na.rm = TRUE)),
    .groups       = "drop"
  )%>%
  mutate(
    geo_level = "county"
  )



cnty_adj_df_all <- rbind(cnty_adj_case_df, cnty_adj_race_df) %>%
  ungroup()

all_rates_dem_adj <- all_rates_by_demographic %>%
  left_join(cnty_adj_df_all, by = c("county", "geo_level", "group_var")) %>%
  left_join(cnty_joint_adj, by = c("county", "geo_level"))



### -- some rigging here to join my sex-age adjusted rates to the dataframe
join_keys <- c(
  "health_officer_region",
  "county",
  "group_var_cat",  
  "group_var",     
  "geo_level"       
)

joined <- all_rates_dem_adj %>%
  left_join(
    cnty_sex_age_adj %>%
      select(all_of(join_keys), adj_sev_100k, adj_rate_100k),
    by = join_keys,
    suffix = c("", "_new")
  )


joined <- joined %>%
  mutate(
    adj_sev_100k  = dplyr::coalesce(adj_sev_100k_new,  adj_sev_100k),
    adj_rate_100k = dplyr::coalesce(adj_rate_100k_new, adj_rate_100k)
  ) %>%
  select(
    -adj_sev_100k_new,
    -adj_rate_100k_new
  )

##########################################################
## Export / overwrite final to 'data/cleaned_data/' dir ##
##########################################################

#write.csv(all_rates_dem_adj, here("data/cleaned_data/all_rates_dem_adj.csv"), row.names = FALSE)
#write.csv(joined, here("data/cleaned_data/all_rates_dem_adj_wsex.csv"), row.names = FALSE)


```


