# Course of the Outbreak

```{r, echo = FALSE}

source("ca_mnth_tbl_code.R")


```


```{r, echo = FALSE}


p1_df <- ca_weekly_cases %>% 
  arrange(mmwr_week) %>%
  mutate(
    p_chng = (cumulative_infected - lag(cumulative_infected)) / lag(cumulative_infected) * 100,
    s_dt  = format(as_date(start_date), format = c("%b %d")),
    e_dt  = format(as_date(end_date), format = c("%b %d")),
    cum_lbl = scales::comma(cumulative_infected, big.mark = ",")
  )

max_chng <- max(p1_df$p_chng, na.rm = TRUE)

p1_df <- p1_df %>%
 mutate(   
    p_chng_scaled = case_when(
      is.na(p_chng) ~ 4,
      max(p_chng, na.rm = TRUE) == min(p_chng, na.rm = TRUE) ~ 8,
      TRUE ~ scales::rescale(
        p_chng,
        to   = c(4, 14),
        from = range(p_chng, na.rm = TRUE)
      )
    )
  )


p1 <- 
plot_ly(
  data = p1_df,
  name = "Cumulative Infections",
  x = ~mmwr_week,
  y = ~cumulative_infected,
  type = "scatter",
  mode = "lines",
  line = list(color = "#fbc9d0")
)

p1 <- p1 %>%
  
add_markers(
    name = "% Change New Cases",
    x = ~mmwr_week,
    y = ~cumulative_infected,
    size =  ~p_chng,
    hoverinfo = "text",
    text = ~paste0(
        "<b>", s_dt, " - ", e_dt, "</b>",
        "<br>Cumulative infected: <b>", cum_lbl, "</b>",
        "<br>Percent change: ", "<b>", round(p_chng, 1), "% </b>"
        ),
    
    marker = list(
        sizemode = "diameter",
        sizeref = 2 * max_chng / (60),
        sizemin  = 3,
        color = "#fdacb8"
      )
)


p1 <- p1 %>%
  
layout(
    plot_bgcolor = "#0f172a",
    paper_bgcolor = "#0f172a",
    font = list(color = '#f1f0ea'),
  title = list(text = "Statewide Cumulative Infections"),
    
    xaxis = list(title = "MMWR Week"),
    yaxis = list(title = "Cumulative Infections"),
    
    showlegend = TRUE,
    legend = list(orientation = 'h', x = 0, y = 1.1),
  
    margin = list(
    l = 60,   
    r = 40,   
    t = 100,  
    b = 60,
    pad = 4 
  )
    
)

```




```{r, echo = FALSE}



p2_df <- p1_df %>%
  select(county, mmwr_week, start_date, end_date, 
         new_infections, new_severe) %>%
  mutate(
    all_inf = new_infections + new_severe,
    inf_prop = 100*new_infections/all_inf,
    sev_prop = 100*new_severe/all_inf,
    across(.cols = where(is.numeric), .fns  = ~ round(.x))
    
  ) %>%
  
  tidyr::pivot_longer(
    cols      = -c(county, mmwr_week, start_date, end_date, all_inf),
    names_to  = "inf_type",
    values_to = "inf_cnt"
  ) 

p2 <- 
ggplot() + 
  geom_col(
    data = p2_df %>% filter(!inf_type %in% c("inf_prop", "sev_prop")), 
    aes(
      x = mmwr_week, 
      y = inf_cnt,
      fill = inf_type
    )
  )  


p3 <- 
ggplot() + 
  geom_col(
    data = p2_df %>% filter(inf_type %in% c("inf_prop", "sev_prop")), 
    aes(
      x = mmwr_week, 
      y = inf_cnt,
      fill = inf_type
    )
  ) 

```




|
|



## Cumulative Cases by Week in California
##### *May 2023 - Dec 2023*

```{r, echo = FALSE}

p1

```


|



## Timeline of the California Outbreak:

|


#### *May 2023 - December 2023*


::: {.indent-block}

##### *May 28, 2023:*

California had reported 13,875 infections. Infections increased rapidly throughout June and early July, with weekly new cases rising **50 to 100 percent** compared with the previous week, the most instense growth phase of the outbreak this year.


##### *Late July - September:*

The outbreak began to slow. Weekly increases in new cases dropped to **20 to 30 percent.**


##### *October:*

Growth had slowed even further, with weekly increases approaching single digits.

##### *Mid-November:* 

Growth in weekly new cases slowed to less than 1%


##### *As of December 30:* 

Although overall infections have continued to decline, we are still getting reports of new cases. 

Most importantly is that there is now a much **higher proportion of new infections that are severe and require hospitalization.**

While the State of California is seeing dramatically fewer infections overall, seeing a higher proportion of infections that are severe requires our continued vigilance and dedication to address this outbreak as a community: maintaining clinical capacity, strengthening rapid detection, and ensuring that the communities being hit the hardest get the resources they need. 

:::



|
|


::: {.wide-embed}

```{r, echo = FALSE}
#| layout-nrow: 1

p2
p3

```

:::


|
|



## Statewide Infections by Month
##### *May 2023 - Dec 2023*


```{r, echo = FALSE}

ca_mnth_tbl

```



|
|


|
|


#### Supporting Code: 

|


[Download Full Table Code (ZIP)](../../www/ca_mnth_tbl_code.zip){.btn .btn-primary target="_blank"}


|
|


**Cumulative Infections Plotly:**

```{r, eval = FALSE}


p1_df <- ca_weekly_cases %>% 
  arrange(mmwr_week) %>%
  mutate(
    p_chng = (cumulative_infected - lag(cumulative_infected)) / lag(cumulative_infected) * 100,
    s_dt  = format(as_date(start_date), format = c("%b %d")),
    e_dt  = format(as_date(end_date), format = c("%b %d")),
    cum_lbl = scales::comma(cumulative_infected, big.mark = ",")
  )

max_chng <- max(p1_df$p_chng, na.rm = TRUE)

p1_df <- p1_df %>%
 mutate(   
    p_chng_scaled = case_when(
      is.na(p_chng) ~ 4,
      max(p_chng, na.rm = TRUE) == min(p_chng, na.rm = TRUE) ~ 8,
      TRUE ~ scales::rescale(
        p_chng,
        to   = c(4, 14),
        from = range(p_chng, na.rm = TRUE)
      )
    )
  )


p1 <- 
plot_ly(
  data = p1_df,
  name = "Cumulative Infections",
  x = ~mmwr_week,
  y = ~cumulative_infected,
  type = "scatter",
  mode = "lines",
  line = list(color = "#fbc9d0")
)

p1 <- p1 %>%
  
add_markers(
    name = "% Change New Cases",
    x = ~mmwr_week,
    y = ~cumulative_infected,
    size =  ~p_chng,
    hoverinfo = "text",
    text = ~paste0(
        "<b>", s_dt, " - ", e_dt, "</b>",
        "<br>Cumulative infected: <b>", cum_lbl, "</b>",
        "<br>Percent change: ", "<b>", round(p_chng, 1), "% </b>"
        ),
    
    marker = list(
        sizemode = "diameter",
        sizeref = 2 * max_chng / (60),
        sizemin  = 3,
        color = "#fdacb8"
      )
)


p1 <- p1 %>%
  
layout(
    plot_bgcolor = "#0f172a",
    paper_bgcolor = "#0f172a",
    font = list(color = '#f1f0ea'),
  title = list(text = "Statewide Cumulative Infections"),
    
    xaxis = list(title = "MMWR Week"),
    yaxis = list(title = "Cumulative Infections"),
    
    showlegend = TRUE,
    legend = list(orientation = 'h', x = 0, y = 1.1),
  
    margin = list(
    l = 60,   
    r = 40,   
    t = 100,  
    b = 60,
    pad = 4 
  )
    
) 

```


|


**Final table build:**

```{r, eval = FALSE}


###########################################
###########################################
##                                       ##
##         Build the Reactable           ##
##                                       ##
###########################################
###########################################


ca_mnth_tbl <- 
  
reactable::reactable(
  data,
  pagination = FALSE,
  defaultSorted    = "metric",
  defaultSortOrder = "asc",
  class = "reactable",
    
  defaultColDef = reactable::colDef(
    vAlign = "center",
    headerVAlign = "bottom",
    sortable    = FALSE 
  ),
    
  columnGroups = list(
    
    reactable::colGroup(
      name = "Outbreak by Month",
      columns = mnth_cols,
    )
  ),
    
  columns = c(
      list(
        metric = reactable::colDef(
          name = "Metric",
          align = "left",
          maxWidth = 80
        )
        
      ),
      
      month_cols
    )
 
####################  
##-- end of table 
)


```



