---
echo: false
---


# Course of the Outbreak

::: {.vspace}

:::


```{r, echo = FALSE}

source("ca_mnth_tbl_code.R")


```



::: {.vspace}

:::


*Please hang tight. The application below may take a minute or two to load.*

::: {.vspace}

:::


::: {.wide-embed}

```{=html}

<iframe 
  src="https://evalue.shinyapps.io/Group08_County_Weekly_Map/"
  width="100%" 
  height="1000px" 
  frameborder="0">
</iframe>

```






::: {.vspace-lg}

:::



::: {.hr-colored .purple}

:::



## Statewide Infections by Month
##### *May 2023 - Dec 2023*


```{r, echo = FALSE}

ca_mnth_tbl

```


::: {.vspace}

:::


::: {.small-note-nopad}

* For each metric, cell shading reflects the **relative contribution of each month to the total across all months combined**. 

* Darker shading = months with higher infection counts or rates relative to other months

* Infection and Severe Infection rates are reported as the number of cases per 100,000 population

:::




::: {.vspace-lg}

:::



## Timeline of the Outbreak

### *May 2023 - December 2023*

--------------

::: {.indent-block}

##### *May 28, 2023:* 

Cumulative Infections: **13,875** 

::: {.vspace}

:::

##### *June:* 

Cumulative Infections: **133,835**

Infections increased rapidly, with weekly new cases rising **50 to 100 percent** compared with the previous week. This was the most intense growth phase of the outbreak. By July 8th, we saw a 864.58% increase since May.

::: {.vspace}

:::

##### *July:* 

Cumulative Infections: **441,588**

The outbreak began to slow in terms of it's week-over percent growth, dropping to about **20 to 30 percent,** but the number of new infections reported each week continued to rise. By the end of July, there was a 229.95% increase.  

::: {.vspace}

:::


```{r}
sepoct <- ca_weekly_cases %>%
  filter(mmwr_week %in% c(36:43)) 
```


##### *September-October:* 

Cumulative Infections: **4,185,531**

We saw the largest growth in cumulative infections statewide during September and October, with an average of over 334,000 new cases reported each week. 

::: {.vspace}

:::

##### *November:* 

Cumulative Infections: **4,533,887**

Growth in weekly percent change in new cases drops precipitously to less than 1%, and we start to see the outbreak slow down. 

::: {.vspace}

:::


##### *December 30:* 

Cumulative Infections: **4,549,583**

New infections have continued to decline each week, but we are still getting reports of new cases. 

Most importantly is that there is now a much **higher proportion of new infections that are severe and require hospitalization.**

While the State of California is seeing dramatically fewer infections overall, seeing a higher proportion of infections that are severe requires our continued vigilance and dedication to address this outbreak as a community: maintaining clinical capacity, strengthening rapid detection, and ensuring that the communities being hit the hardest get the resources they need. 

:::


::: {.vspace-lg}

:::




::: {.hr-colored .purple}

:::




:::


