# Course of the Outbreak

::: {.vspace}

:::


```{r, echo = FALSE}

source("ca_mnth_tbl_code.R")


```



::: {.vspace}

:::


*Please hang tight. The application below may take a minute or two to load.*

::: {.vspace}

:::


::: {.wide-embed}

```{=html}

<iframe 
  src="https://evalue.shinyapps.io/Group08_County_Weekly_Map/"
  width="100%" 
  height="1000px" 
  frameborder="0">
</iframe>

```






::: {.vspace-lg}

:::



::: {.hr-colored .purple}

:::



## Statewide Infections by Month
##### *May 2023 - Dec 2023*


```{r, echo = FALSE}

ca_mnth_tbl

```




::: {.vspace-lg}

:::




## Timeline of the Outbreak

### *May 2023 - December 2023*

--------------

::: {.indent-block}
##### *May 28, 2023:*

California had reported 13,875 infections. Infections increased rapidly throughout June and early July, with weekly new cases rising **50 to 100 percent** compared with the previous week, the most instense growth phase of the outbreak this year.


##### *Late July - September:*

The outbreak began to slow. Weekly increases in new cases dropped to **20 to 30 percent.**


##### *October:*

Growth had slowed even further, with weekly increases approaching single digits.

##### *Mid-November:* 

Growth in weekly percent change in new case slowed to less than 1%


##### *As of December 30:* 

Although overall infections have continued to decline, we are still getting reports of new cases. 

Most importantly is that there is now a much **higher proportion of new infections that are severe and require hospitalization.**

While the State of California is seeing dramatically fewer infections overall, seeing a higher proportion of infections that are severe requires our continued vigilance and dedication to address this outbreak as a community: maintaining clinical capacity, strengthening rapid detection, and ensuring that the communities being hit the hardest get the resources they need. 

:::


::: {.vspace-lg}

:::




::: {.hr-colored .purple}

:::




:::


