---
echo: false
---


```{r}
#| include: false
#| warning: false
#| message: false

source("_common04.R")
source("04bb_race_code.R")

```



# Burden of Disease: Race


::: {.vspace}

:::

::: {.vspace}

:::


::: {.full-bleed-banner}
::: {.banner-inner}


::: {.vspace}

:::

::: {.center-text}

##### Proportion of CA Population by Race


```{r, fig.height=4}
race_prop_pie
```


:::

::: {.vspace}

:::


:::: {.columns}

::: {.column width="33%" .exsmall-note}

**White NH** = White, Non-Hispanic  

**Black NH** = Black, Non-Hispanic  

**AIAN NH** = American Indian or Alaska Native, Non-Hispanic

:::

::: {.column width="33%" .exsmall-note}

**Asian NH** = Asian, Non-Hispanic 

**NHPI NH** = Native Hawaiian or Pacific Islander, Non-Hispanic  

**MR NH** = Multiracial (two or more of the above races), Non-Hispanic

:::

::: {.column width="33%" .exsmall-note}

**Hispanic** = Hispanic (any race)  

**Unknown** = Unknown

:::

::::

::: {.vspace}

:::


:::
:::





::: {.vspace}

:::


### Infections by Race / Ethnicity

::: {.vspace}

:::


Interpreting racial differences in infection rates is complex because observed patterns often reflect multiple overlapping social and structural factors. 


::: {.vspace}

:::



::: {.indent-block}

Adjusting for both age and race is a way for us to ask: 

*"If all counties had the same age structure and racial composition, what would infection rates look like?"* 
  
:::


::: {.vspace}

:::


Race-adjusted rates help reduce the influence of differing racial compositions across counties, but they cannot explain *why* a population group shows elevated risk. Higher rates may reflect individual-level vulnerabilities, structural inequities such as racism or unequal access to resources, county-level environmental or social conditions, or the interaction among these factors.  



For example,  after adjusting for both age and race across counties:


::: {.circle-bullet .purple}

Counties with the largest **Hispanic populations** experienced the highest rates of  **severe infections** on average.

:::

This finding signals an important disparity, but the adjusted rate alone cannot identify its underlying causes. One possible explanation is that many of **these counties are located near major urban areas with very high numbers of cases, such as Los Angeles, Sacramento, and San Bernardino counties**, which may increase exposure risk. 

Additional analysis would be needed to determine whether this pattern reflects geographic proximity, social determinants of health, barriers to prevention and treatment, or a combination of factors.



::: {.vspace}

:::


::: {.hr-colored .purple}

:::




::: {.vspace}

:::




::: card

#### County Adjusted Infection Rates by Hispanic Population {.card-header}

::: {.card-body}


```{r}
#| fig-width: 8
#| fig-height: 5

race_p1_plotly

```


:::

:::


::: {.vspace}

:::


::: {.small-note}

*The severe infection rates were adjusted jointly for both age distributions and racial composition (reported as rates per 100,000 population).*


:::



::: {.vspace}

:::

::: {.vspace}

:::




::: {.hr-colored .purple}

:::



::: {.vspace-lg}

:::




::: {.center-text}

#### Statewide Proportion of Infections by Race

:::



::: {layout="[ [1], [1,1] ]"}


```{r, fig.height= 4}

race_bar_p

```


::: card

#### Statewide Proportion of Infections by Race {.card-header}

::: {.card-body}

```{r}
race_tbl_1
```

:::
:::


::: card

#### Distribution of Severe Infections Across Counties{.card-header}

::: {.card-body}

```{r}
race_tbl_2
```

:::
:::


:::




:::: {.columns}

::: {.column width="33%" .exsmall-note}

**White NH** = White, Non-Hispanic  

**Black NH** = Black, Non-Hispanic  

**AIAN NH** = American Indian or Alaska Native, Non-Hispanic

:::

::: {.column width="33%" .exsmall-note}

**Asian NH** = Asian, Non-Hispanic 

**NHPI NH** = Native Hawaiian or Pacific Islander, Non-Hispanic  

**MR NH** = Multiracial (two or more of the above races), Non-Hispanic

:::

::: {.column width="33%" .exsmall-note}

**Hispanic** = Hispanic (any race)  

**Unknown** = Unknown

:::

::::




