---
echo: false
---


```{r}
#| include: false
#| warning: false
#| message: false

source("_common04.R")
source("04bb_race_code.R")

```



# Burden of Disease: Race


::: {.vspace}

:::

::: {.vspace}

:::


::: {.full-bleed-banner}
::: {.banner-inner}


::: {.vspace}

:::

::: {.center-text}

##### Proportion of CA Population by Race


```{r, fig.height=4}
race_prop_pie
```


:::

::: {.vspace}

:::


:::: {.columns}

::: {.column width="33%" .exsmall-note}

**White NH** = White, Non-Hispanic  

**Black NH** = Black, Non-Hispanic  

**AIAN NH** = American Indian or Alaska Native, Non-Hispanic

:::

::: {.column width="33%" .exsmall-note}

**Asian NH** = Asian, Non-Hispanic 

**NHPI NH** = Native Hawaiian or Pacific Islander, Non-Hispanic  

**MR NH** = Multiracial (two or more of the above races), Non-Hispanic

:::

::: {.column width="33%" .exsmall-note}

**Hispanic** = Hispanic (any race)  

**Unknown** = Unknown

:::

::::

::: {.vspace}

:::


:::
:::





::: {.vspace}

:::


### Infections by Race / Ethnicity

::: {.vspace}

:::


Interpreting racial differences in infection rates is complex because observed patterns often reflect multiple overlapping social and structural factors. 


::: {.vspace}

:::



::: {.indent-block}

Adjusting for both age and race is a way for us to ask: 

*"If all counties had the same age structure and racial composition, what would infection rates look like?"* 
  
:::


::: {.vspace}

:::


Race-adjusted rates help reduce the influence of differing racial compositions across counties, but they cannot determine *why* a population groups shows elevated risk: whether it reflects individual-level vulnerabilities, county-level conditions, or the interaction of both. It is important to recognize that race is a social construct shaped by historical and structural forces. Because of this, race can serve as a proxy for inequities rather than a direct cause of disease. In some cases, statistical adjustment may unintentionally obscure meaningful disparities that stem from structural racism and unequal access to resources. 



However, even after adjusting for both age and race across counties, one key finding persisted:


::: {.circle-bullet .purple}

Counties with the largest **Hispanic populations** experienced the highest rates of  **severe infections** on average.

:::





::: {.wide-embed}


::: {.hr-colored .purple}

:::


::: {.center-text}

#### Adjusted Severe Infection Rates and Proportion of Population by County

::: 



::: {.vspace}

:::


::: {layout="[[1.5,1]]"}




::: card

#### Adjusted Infection Rates by County Hispanic Population Proportion {.card-header}

::: {.card-body}


```{r}
#| fig-width: 7
#| fig-height: 4.5

race_p1_plotly

```


:::
:::


::: card

#### Adjusted Infection Rates by all other Races {.card-header}

::: {.card-body}

```{r}
#| fig-width: 8
#| fig-height: 7


race_p2

```


:::
:::


:::


::: {.small-note}

*The severe infection rates were adjusted jointly for both age distributions and racial composition (reported as rates per 100,000 population).*

*The line shows the average trend across counties and the band shows how much the estimate could vary due to uncertainty.*

:::



::: {.vspace-xlg}

:::




::: {.hr-colored .purple}

:::



::: {.vspace}

:::








::: {.center-text}

#### Statewide Proportion of Infections by Race

:::



::: {layout="[ [1], [1,1] ]"}


```{r, fig.height= 4}

race_bar_p

```


::: card

#### Statewide Proportion of Infections by Race {.card-header}

::: {.card-body}

```{r}
race_tbl_1
```

:::
:::


::: card

#### Distribution of Severe Infections Across Counties{.card-header}

::: {.card-body}

```{r}
race_tbl_2
```

:::
:::


:::




:::: {.columns}

::: {.column width="33%" .exsmall-note}

**White NH** = White, Non-Hispanic  

**Black NH** = Black, Non-Hispanic  

**AIAN NH** = American Indian or Alaska Native, Non-Hispanic

:::

::: {.column width="33%" .exsmall-note}

**Asian NH** = Asian, Non-Hispanic 

**NHPI NH** = Native Hawaiian or Pacific Islander, Non-Hispanic  

**MR NH** = Multiracial (two or more of the above races), Non-Hispanic

:::

::: {.column width="33%" .exsmall-note}

**Hispanic** = Hispanic (any race)  

**Unknown** = Unknown

:::

::::





:::



