# Weekly Infection Map


```{r}
#| include: false
#| warning: false
#| message: false

source("_common04.R")

```


[Download Weekly Infection Rates (CSV)](../../data/shapefiles/cnty_week_pnts.csv){.btn .btn-primary target="_blank"}


|
|



::: {.wide-embed}

::: {.panel-tabset}


## Map

#### Infection Rates by Week
Adjust the slider to follow the progression of the outbreak week by week. Every point reflects the infection rate or severe infection rate for a California county, standardized per 100,000 population. Counties that are shaded are those in the highest top five for either infection or severe infection rates. 



```{=html}

<iframe 
  src="https://evalue.shinyapps.io/Group08_County_Weekly_Map/"
  width="100%" 
  height="740px" 
  frameborder="0">
</iframe>

```

:::

:::


|
|


#### Supporting Code: 

The full code for the interactive maps can be found on the github repository within the "cases_week_app" branch:
**<https://github.com/EvalstacE/Group08_PHW251_R/tree/cases_week_app>**


[Download Shiny App Code (ZIP)](../../www/Group08_weekly_case_map_shiny_app.zip){.btn .btn-primary target="_blank"}



|
|





**Creating and saving shapefiles for maps:**

```{r, eval = FALSE}

##--bring in California shapefile and 
###---join to infection rate dataframe
ca_cnty_sf <- counties(state = "CA", cb = TRUE, year = 2023) %>%
  st_as_sf() %>%
  select(NAME) %>%
  left_join(inf_rates_by_cnty, by = c("NAME" = "county")) %>%
  rename("county" = "NAME") %>%
  select(county, health_officer_region) %>%
  st_transform(4326)

##--state border only
ca_state_sf <- states(cb = TRUE, year = 2023) %>%
  st_as_sf() %>%
  filter(STUSPS == "CA") %>%
  select(STATEFP, STUSPS, GEOIDFQ, GEOID) %>%
  st_transform(4326)


##--dissolves counties into Health Officer Regions
hor_sf <- ca_cnty_sf %>%
  group_by(health_officer_region) %>%
  summarise(.groups = "drop")%>%
  st_cast("MULTIPOLYGON")%>%
  st_transform(4326)

##--creates points centered in polygons
ca_cnty_pnts <- ca_cnty_sf %>%
  st_centroid() %>%              
  mutate(
    lng = st_coordinates(.)[, 1],
    lat = st_coordinates(.)[, 2]
  ) %>%
  st_drop_geometry() %>%
  select(county, lng, lat)


hor_pnts <- hor_sf %>%
  st_centroid() %>%              
  mutate(
    lng = st_coordinates(.)[, 1],
    lat = st_coordinates(.)[, 2]
  ) %>%
  st_drop_geometry() %>%
  select(health_officer_region, lng, lat)


##--save to data/shapefiles folder
##-st_write(ca_cnty_sf, here("data/shapefiles/ca_cnty_sf/ca_cnty_sf.shp"), append = FALSE)
##-st_write(hor_sf, here("data/shapefiles/hor_sf/hor_sf.shp"), append = FALSE)
##-st_write(ca_state_sf, here("data/shapefiles/ca_state_sf/ca_state_sf.shp"), append = FALSE)
##-write.csv(ca_cnty_pnts, here("data/weekly_case_points/ca_cnty_pnts.csv"), row.names = FALSE)
##-write.csv(hor_pnts, here("data/weekly_case_points/hor_pnts.csv"), row.names = FALSE)

```



**Create weekly infection and severe case dataframe by county and health officer region:**

```{r, eval = FALSE}

## -- weekly case data by county
cnty_weekly_cases <- combined_df %>%
group_by(
    health_officer_region,
    county,
    mmwr_week,
    start_date,
    end_date
) %>%
summarise(
    cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
    cumulative_severe = sum(cumulative_severe, na.rm = TRUE),
    total_cnty_pop = first(total_cnty_pop), 
    inf_rate_100k = round((10^5 * cumulative_infected / total_cnty_pop),1),
    sev_rate_100k = round((10^5 * cumulative_severe / total_cnty_pop),1),
    .groups = "drop"
) %>%
add_EQ_labels(rename_before = FALSE)


## -- weekly case data by HOR
hor_weekly_cases <- combined_df %>%
group_by(
    health_officer_region,
    mmwr_week,
    start_date,
    end_date
) %>%
summarise(
    cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
    cumulative_severe = sum(cumulative_severe, na.rm = TRUE),
    total_HOR_pop = first(total_HOR_pop), 
    inf_rate_100k = round((10^5 * cumulative_infected / total_HOR_pop),1),
    sev_rate_100k = round((10^5 * cumulative_severe / total_HOR_pop),1),
    .groups = "drop"
) %>%
add_EQ_labels(rename_before = FALSE)


#- Export / save dfs:
##-write.csv(cnty_weekly_cases, here("data/weekly_case_points/cnty_weekly_cases.csv"), row.names = FALSE)
##-write.csv(hor_weekly_cases, here("data/weekly_case_points/hor_weekly_cases.csv"), row.names = FALSE)

```




**Append points to weekly infection data:**

```{r, eval = FALSE}


## -- make weekly point data
cnty_week_pnts <- cnty_weekly_cases %>%
  left_join(ca_cnty_pnts, by = "county")

hor_week_pnts <- hor_weekly_cases %>%
  left_join(hor_pnts, by = "health_officer_region")


##--save to data/shapefiles folder
##-write.csv(cnty_week_pnts, here("data/weekly_case_points/cnty_week_pnts.csv"), row.names = FALSE)
##-write.csv(hor_week_pnts, here("data/weekly_case_points/hor_week_pnts.csv"), row.names = FALSE)

```

