# Infections by Age, Sex, and Race


```{r}
#| include: false
#| warning: false
#| message: false

source("_common04.R")

```


When we look across the entire state, females have a lower overall infection rate than males, but a higher rate of severe infections. This same pattern holds when infection and severe infection rates are compared within each health officer region.

The table below shows infection and severe infection rates by sex and by health officer region.


#### Infection and Severe Infection Rates by Sex and Health Officer Region
##### *California Outbreak, May 2023 - Dec 2023*

```{r, eval = FALSE}

##########################
#--Prepare dataframe
##--see 04i_calc_rates_rng_grp.R function
###--in 'functions/' folder

# specs: which grouping var uses which population column
group_specs <- tibble::tibble(
  group_var = list(quo(sex), quo(age_cat), quo(race_short)),
  pop_col   = list(quo(total_sex_pop), quo(pop), quo(total_race_pop))
)

############################
##  Region-level (HOR)   ###
############################

hor_results_list <- group_specs %>%
  purrr::pmap(function(group_var, pop_col) {
    calc_rates_rng_grp(
      df        = combined_df,
      group_var = !!group_var,
      pop_col   = !!pop_col,
      geo_level = "region"   
    ) %>%
      dplyr::mutate(geo_level = "region")
  })

hor_rates_by_demographic <- dplyr::bind_rows(hor_results_list)

############################
##  County-level         ###
############################

cnty_results_list <- group_specs %>%
  purrr::pmap(function(group_var, pop_col) {
    calc_rates_rng_grp(
      df        = combined_df,
      group_var = !!group_var,
      pop_col   = !!pop_col,
      geo_level = "county"
    ) %>%
      dplyr::mutate(geo_level = "county")
  })

cnty_rates_by_demographic <- dplyr::bind_rows(cnty_results_list)

############################################
##   all rates by dem combined in one df  ##
############################################

all_rates_by_demographic <- dplyr::bind_rows(
  hor_rates_by_demographic,
  cnty_rates_by_demographic
)

  
#########################################
## Exports to 'data/cleaned_data/' dir ##
#########################################

#write.csv(cnty_rates_by_demographic, here("data/cleaned_data/cnty_rates_by_demographic.csv"), row.names = FALSE)
#write.csv(hor_rates_by_demographic, here("data/cleaned_data/hor_rates_by_demographic.csv"), row.names = FALSE)
#write.csv(all_rates_by_demographic, here("data/cleaned_data/all_rates_by_demographic.csv"), row.names = FALSE)

```





```{r, echo = FALSE}

##########################
#--Build table for display

sex_tbl_df <- by_sex %>%
  select(-c(cumulative_infected, cumulative_severe)) %>%
  rename(
    `Health Officer Region`  = health_officer_region,
    Sex                      = sex,
    Population               = total_sex_pop,
    `Infection Rate (/100k)` = inf_rate_100k,
    `Severe Rate (/100k)`    = sev_rate_100k
  ) 

sex_df_disp_grp <- sex_tbl_df %>%
  group_by(`Health Officer Region`) %>%
  mutate(
    `Health Officer Region` = if_else(row_number() == 1, `Health Officer Region`, "")
  ) %>%
  ungroup()



n_rows <- nrow(sex_df_disp_grp)
shade_rows <- unlist(
  mapply(`:`,
         seq(3, n_rows, by = 4),
         seq(4, n_rows, by = 4))
)


sex_tbl <- sex_df_disp_grp %>%
kbl(
  align   = c("l", "l", "c", "c", "c"),
  format  = "html", 
  escape  = FALSE
) %>%
  
row_spec(
  0,
  bold = TRUE,
  background = "#0f172a",
  extra_css = "font-size: 16px!important;color:#ffffff;"
)%>%
  
column_spec(
  5,
  bold = TRUE,
  extra_css = "font-size: 14px!important;color:#0f172a;"
)%>%
  
row_spec(
  seq(2, n_rows, by = 2),
  extra_css = "border-bottom: 1.25px solid #0f172a;"
) %>%
  
row_spec(
    shade_rows,
    background = "#f1f0ea"
) %>%  
  
kable_styling(bootstrap_options = c("hover"), full_width = TRUE)

```


::: {.no-stripe}


```{r, echo = FALSE}

sex_tbl

```

:::



|
|
|
|

#### Supporting Code: 

**Code for table:**

| 


```{r, eval = FALSE}

by_sex1 <- combined_df %>%
  #--only last week (52)
  filter(mmwr_week == 52) %>%
  
  #--one row per region–county–sex 
  group_by(health_officer_region, county, sex) %>%
  summarise(
    total_sex_pop       = first(total_sex_pop),
    cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
    cumulative_severe   = sum(cumulative_severe, na.rm = TRUE),
    .groups = "drop"
  ) 

by_sex <- by_sex1 %>%  
  #--sum county pops and infectionswithin each region–sex
  group_by(health_officer_region, sex) %>%
  summarise(
    total_sex_pop       = sum(total_sex_pop, na.rm = TRUE),
    cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
    cumulative_severe   = sum(cumulative_severe, na.rm = TRUE),
    inf_rate_100k       = round((10^5*cumulative_infected/total_sex_pop)),
    sev_rate_100k       = round((10^5*cumulative_severe/total_sex_pop)),
    .groups = "drop"
  )
  

sex_tbl_df <- by_sex %>%
  select(-c(cumulative_infected, cumulative_severe)) %>%
  rename(
    `Health Officer Region`  = health_officer_region,
    Sex                      = sex,
    Population               = total_sex_pop,
    `Infection Rate (/100k)` = inf_rate_100k,
    `Severe Rate (/100k)`    = sev_rate_100k
  ) 

sex_df_disp_grp <- sex_tbl_df %>%
  group_by(`Health Officer Region`) %>%
  mutate(
    `Health Officer Region` = if_else(row_number() == 1, `Health Officer Region`, "")
  ) %>%
  ungroup()



n_rows <- nrow(sex_df_disp_grp)
shade_rows <- unlist(
  mapply(`:`,
         seq(3, n_rows, by = 4),
         seq(4, n_rows, by = 4))
)


sex_tbl <- sex_df_disp_grp %>%
kbl(
  align   = c("l", "l", "c", "c", "c"),
  format  = "html", 
  escape  = FALSE
) %>%
  
row_spec(
  0,
  bold = TRUE,
  background = "#0f172a",
  extra_css = "font-size: 16px!important;color:#ffffff;"
)%>%
  
column_spec(
  5,
  bold = TRUE,
  extra_css = "font-size: 14px!important;color:#0f172a;"
)%>%
  
row_spec(
  seq(2, n_rows, by = 2),
  extra_css = "border-bottom: 1.25px solid #0f172a;"
) %>%
  
row_spec(
    shade_rows,
    background = "#f1f0ea"
) %>%  
  
kable_styling(bootstrap_options = c("hover"), full_width = TRUE)

```

