---
echo: false
---



# Resource Prioritization

::: {.vspace}

:::


```{r}
#| include: false
#| warning: false
#| message: false

source("_common04.R")
source("cum_cases_map.R")

```


### Recommended Prioritization Strategy

Based on the statewide analysis of infection rates, severe disease burden, and population distribution, we recommend a two-part strategy to guide allocation of prevention, testing, and treatment resources. 

This approach recognizes that total case counts and per capita rates each reflect different types of vulnerability. Urban counties require strong prevention and testing to reduce rapid transmission at scale. Rural counties need enhanced treatment and recovery support because their smaller populations experience a much higher proportional disease burden. Prioritizing resources based on both absolute and relative impacts provides a balanced and equitable strategy that reduces statewide transmission while protecting communities experiencing the most severe outcomes.


::: {.highlight-text}
::: {.circle-bullet .purple}

**1. Prioritize prevention and testing capacity in densely populated urban counties with rapid transmission growth.**

:::
:::

::: {.vspace}

:::

##### Los Angeles, San Bernardino, and Sacramento Counties

Reported the largest number of total infections and demonstrate the fastest acceleration in weekly transmission. High population density increases the likelihood of widespread community spread, which in turn elevates risk for neighboring counties and the state as a whole. Concentrating early prevention and control activities in these settings is likely to have the greatest effect on slowing statewide transmission. This could include expanding public education and outreach, increasing access to low-barrier testing, and supporting local public health policies that help control disease spread. 



::: {.highlight-text}
::: {.circle-bullet .purple}

**2. Prioritize treatment, clinical support, and recovery resources in counties with the highest per capita burden**

:::
:::



::: {.vspace}

:::

##### Imperial, Inyo, Kings, Tulare, Kern, and Merced Counties 

Experienced the highest infection and severe infection rates in California, despite having relatively small total case counts. These rural counties experience disproportionate impacts because even modest increases in infections translate into very high rates per capita. Directing treatment resources to these areas supports timely clinical care, hospital surge planning, and rapid linkage to treatment. This prioritization helps offset structural disadvantages related to geography, healthcare access, and smaller public health workforces.


::: {.vspace}

:::



**As resources allow, extend priority into second-highest per capita tier:**

##### Colusa, Stanislaus, San Joaquin, Tehama, Sutter, Yuba, and Solano Counties



::: {.vspace}

:::



:::{.wide-embed}



::: {layout="[ [1,1] ]"}

::: card

#### Cumulative Infections by County {.card-header}


::: {.card-body}

```{r}
#| fig-height: 2.05
#| fig-width: 2


cum_cases_map
```

:::

:::


::: card

#### Age-Adjusted Severe Infection Rate by County {.card-header}

::: {.card-body}

```{r}
#| fig-height: 2.05
#| fig-width: 2

cnty_rank_map
```


:::

:::


:::








::: {.full-bleed-banner}

::: {.banner-inner}

::: {.vspace}

:::


##### Methods for ranking counties based on age-adjusted infection and severe infection rates:

To identify counties with the greatest burden of infection and severe infection, we developed a ranking system based on age-adjusted infection rates and age-adjusted severe infection rates. This approach highlights counties experiencing consistently elevated disease burden and allows us to identify and target the areas with the greatest need for additional support and resources.

We first calculated statewide percentiles for each age-adjusted rate. Counties were compared against the 90th, 75th, and 50th percentile cut points for both infection and severe infection rates. Counties with rates at or above the 90th percentile were assigned a rank of 1, those between the 75th and 90th percentiles received a rank of 2, those between the 50th and 75th percentiles received a rank of 3, and counties below the 50th percentile received a rank of 4.

Each county received two percentile-based ranks, one for its adjusted infection rate and one for its adjusted severe infection rate. These two values were summed to create an overall priority score, which reflects the combined burden of disease and severe disease at the county level.

* **Top Priority (score of 2):** Both the age-adjusted infection rate and the age-adjusted severe infection rate for the county fall at or above the 90th percentile statewide.

* **Second Priority (score of 3 or 4):** A county is classified as second priority if either one rate is at or above the 90th percentile and the other is at or above the 75th percentile, or if both rates fall within the 75th percentile range


::: {.vspace}

:::

:::

:::




```{=html}

<iframe 
  src="https://evalue.shinyapps.io/Group08_cases_dem_app/"
  width="100%" 
  height="2000px" 
  frameborder="0">
</iframe>

```



:::




