# Infections by Age and Sex


```{r}
#| include: false
#| warning: false
#| message: false

source("_common04.R")

```


When we looked at infection and severe infection rates across the state of California as a whole, females had a lower infection rate but a higher severe infection rate compared to males. 

When we compared male and female infection rates by health officer region, this trend still appears. 

Below is a table of infections and severe infections broken down by sex and health officer region:


#### Infection and Severe Infection Rates by Sex and Health Officer Region
##### *California Outbreak, May 2023 - Dec 2023*

```{r, echo = FALSE}


by_sex <- combined_df %>%
  group_by(health_officer_region, county, sex) %>%
    summarise(
      total_sex_pop       = first(total_sex_pop),
      cumulative_infected = max(cumulative_infected, na.rm = TRUE),
      cumulative_severe   = max(cumulative_severe, na.rm = TRUE),
      inf_rate_100k       = round((10^5*cumulative_infected/total_sex_pop)),
      sev_rate_100k       = round((10^5*cumulative_severe/total_sex_pop)),
      .groups = "drop"
    ) %>% 
  group_by(health_officer_region, sex) %>%
    summarise(
        total_pop           = sum(total_sex_pop),
        cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
        cumulative_severe   = sum(cumulative_severe, na.rm = TRUE),
        inf_rate_100k       = round((10^5*cumulative_infected/total_pop)),
        sev_rate_100k       = round((10^5*cumulative_severe/total_pop)),
        .groups = "drop"
    ) %>%
  arrange(health_officer_region, sex)
  

sex_tbl_df <- by_sex %>%
  rename(
    `Health Officer Region` = health_officer_region,
    Sex                     = sex,
    Population              = total_pop,
    Infections              = cumulative_infected,
    `Severe Infections`     = cumulative_severe,
    `Infection Rate (/100k)`       = inf_rate_100k,
    `Severe Rate (/100k)`    = sev_rate_100k
  ) 

sex_df_disp_grp <- sex_tbl_df %>%
  group_by(`Health Officer Region`) %>%
  mutate(
    `Health Officer Region` = if_else(row_number() == 1, `Health Officer Region`, "")
  ) %>%
  ungroup()



n_rows <- nrow(sex_df_disp_grp)
shade_rows <- unlist(
  mapply(`:`,
         seq(3, n_rows, by = 4),
         seq(4, n_rows, by = 4))
)


sex_tbl <- sex_df_disp_grp %>%
kbl(
  caption = "Infection and severe infection rates by sex and Health Officer Region",
  align   = c("l", "l", "r", "r", "r", "r", "r"),
  format  = "html", 
  escape  = FALSE
) %>%
  
row_spec(
  0,
  bold = TRUE,
  background = "#0f172a",
  extra_css = "font-size: 16px!important;color:#ffffff;"
)%>%
  
column_spec(
  7,
  bold = TRUE,
  extra_css = "font-size: 14px!important;color:#0f172a;"
)%>%
  
row_spec(
  seq(2, n_rows, by = 2),
  extra_css = "border-bottom: 1.25px solid #0f172a;"
) %>%
  
row_spec(
    shade_rows,
    background = "#f1f0ea"
) %>%  
  
kable_styling(bootstrap_options = c("hover"), full_width = TRUE)


sex_tbl

```




|
|
|
|

#### Supporting Code: 

**Code for table:**

| 


```{r, eval = FALSE}

by_sex <- combined_df %>%
  group_by(health_officer_region, county, sex) %>%
    summarise(
      total_sex_pop       = first(total_sex_pop),
      cumulative_infected = max(cumulative_infected, na.rm = TRUE),
      cumulative_severe   = max(cumulative_severe, na.rm = TRUE),
      inf_rate_100k       = round((10^5*cumulative_infected/total_sex_pop)),
      sev_rate_100k       = round((10^5*cumulative_severe/total_sex_pop)),
      .groups = "drop"
    ) %>% 
  group_by(health_officer_region, sex) %>%
    summarise(
        total_pop           = sum(total_sex_pop),
        cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
        cumulative_severe   = sum(cumulative_severe, na.rm = TRUE),
        inf_rate_100k       = round((10^5*cumulative_infected/total_pop)),
        sev_rate_100k       = round((10^5*cumulative_severe/total_pop)),
        .groups = "drop"
    ) %>%
  arrange(health_officer_region, sex)
  

sex_tbl_df <- by_sex %>%
  rename(
    `Health Officer Region` = health_officer_region,
    Sex                     = sex,
    Population              = total_pop,
    Infections              = cumulative_infected,
    `Severe Infections`     = cumulative_severe,
    `Infection Rate (/100k)`       = inf_rate_100k,
    `Severe Rate (/100k)`    = sev_rate_100k
  ) 

sex_df_disp_grp <- sex_tbl_df %>%
  group_by(`Health Officer Region`) %>%
  mutate(
    `Health Officer Region` = if_else(row_number() == 1, `Health Officer Region`, "")
  ) %>%
  ungroup()



n_rows <- nrow(sex_df_disp_grp)
shade_rows <- unlist(
  mapply(`:`,
         seq(3, n_rows, by = 4),
         seq(4, n_rows, by = 4))
)


sex_tbl <- sex_df_disp_grp %>%
kbl(
  caption = "Infection and severe infection rates by sex and Health Officer Region",
  align   = c("l", "l", "r", "r", "r", "r", "r"),
  format  = "html", 
  escape  = FALSE
) %>%
  
row_spec(
  0,
  bold = TRUE,
  background = "#0f172a",
  extra_css = "font-size: 16px!important;color:#ffffff;"
)%>%
  
column_spec(
  7,
  bold = TRUE,
  extra_css = "font-size: 14px!important;color:#0f172a;"
)%>%
  
row_spec(
  seq(2, n_rows, by = 2),
  extra_css = "border-bottom: 1.25px solid #0f172a;"
) %>%
  
row_spec(
    shade_rows,
    background = "#f1f0ea"
) %>%  
  
kable_styling(bootstrap_options = c("hover"), full_width = TRUE)

```