# Infections by Age and Sex


```{r}
#| include: false
#| warning: false
#| message: false

source("_common04.R")

```


When we look across the entire state, females have a lower overall infection rate than males, but a higher rate of severe infections. This same pattern holds when infection and severe infection rates are compared within each health officer region.

The table below shows infection and severe infection rates by sex and by health officer region.


#### Infection and Severe Infection Rates by Sex and Health Officer Region
##### *California Outbreak, May 2023 - Dec 2023*

```{r, echo = FALSE}


by_sex <- combined_df %>%
  group_by(health_officer_region, county, sex) %>%
    summarise(
      total_sex_pop       = first(total_sex_pop),
      cumulative_infected = max(cumulative_infected, na.rm = TRUE),
      cumulative_severe   = max(cumulative_severe, na.rm = TRUE),
      inf_rate_100k       = round((10^5*cumulative_infected/total_sex_pop)),
      sev_rate_100k       = round((10^5*cumulative_severe/total_sex_pop)),
      .groups = "drop"
    ) %>% 
  group_by(health_officer_region, sex) %>%
    summarise(
        total_pop           = sum(total_sex_pop),
        cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
        cumulative_severe   = sum(cumulative_severe, na.rm = TRUE),
        inf_rate_100k       = round((10^5*cumulative_infected/total_pop)),
        sev_rate_100k       = round((10^5*cumulative_severe/total_pop)),
        .groups = "drop"
    ) %>%
  arrange(health_officer_region, sex)
  

sex_tbl_df <- by_sex %>%
  select(-c(cumulative_infected, cumulative_severe)) %>%
  rename(
    `Health Officer Region` = health_officer_region,
    Sex                     = sex,
    Population              = total_pop,
    `Infection Rate (/100k)` = inf_rate_100k,
    `Severe Rate (/100k)`    = sev_rate_100k
  ) 

sex_df_disp_grp <- sex_tbl_df %>%
  group_by(`Health Officer Region`) %>%
  mutate(
    `Health Officer Region` = if_else(row_number() == 1, `Health Officer Region`, "")
  ) %>%
  ungroup()



n_rows <- nrow(sex_df_disp_grp)
shade_rows <- unlist(
  mapply(`:`,
         seq(3, n_rows, by = 4),
         seq(4, n_rows, by = 4))
)


sex_tbl <- sex_df_disp_grp %>%
kbl(
  align   = c("l", "l", "c", "c", "c"),
  format  = "html", 
  escape  = FALSE
) %>%
  
row_spec(
  0,
  bold = TRUE,
  background = "#0f172a",
  extra_css = "font-size: 16px!important;color:#ffffff;"
)%>%
  
column_spec(
  5,
  bold = TRUE,
  extra_css = "font-size: 14px!important;color:#0f172a;"
)%>%
  
row_spec(
  seq(2, n_rows, by = 2),
  extra_css = "border-bottom: 1.25px solid #0f172a;"
) %>%
  
row_spec(
    shade_rows,
    background = "#f1f0ea"
) %>%  
  
kable_styling(bootstrap_options = c("hover"), full_width = TRUE)

```


::: {.no-stripe}


```{r, echo = FALSE}

sex_tbl

```

:::



|
|
|
|

#### Supporting Code: 

**Code for table:**

| 


```{r, eval = FALSE}

by_sex <- combined_df %>%
  group_by(health_officer_region, county, sex) %>%
    summarise(
      total_sex_pop       = first(total_sex_pop),
      cumulative_infected = max(cumulative_infected, na.rm = TRUE),
      cumulative_severe   = max(cumulative_severe, na.rm = TRUE),
      inf_rate_100k       = round((10^5*cumulative_infected/total_sex_pop)),
      sev_rate_100k       = round((10^5*cumulative_severe/total_sex_pop)),
      .groups = "drop"
    ) %>% 
  group_by(health_officer_region, sex) %>%
    summarise(
        total_pop           = sum(total_sex_pop),
        cumulative_infected = sum(cumulative_infected, na.rm = TRUE),
        cumulative_severe   = sum(cumulative_severe, na.rm = TRUE),
        inf_rate_100k       = round((10^5*cumulative_infected/total_pop)),
        sev_rate_100k       = round((10^5*cumulative_severe/total_pop)),
        .groups = "drop"
    ) %>%
  arrange(health_officer_region, sex)
  

sex_tbl_df <- by_sex %>%
  select(-c(cumulative_infected, cumulative_severe)) %>%
  rename(
    `Health Officer Region` = health_officer_region,
    Sex                     = sex,
    Population              = total_pop,
    `Infection Rate (/100k)`       = inf_rate_100k,
    `Severe Rate (/100k)`    = sev_rate_100k
  ) 

sex_df_disp_grp <- sex_tbl_df %>%
  group_by(`Health Officer Region`) %>%
  mutate(
    `Health Officer Region` = if_else(row_number() == 1, `Health Officer Region`, "")
  ) %>%
  ungroup()



n_rows <- nrow(sex_df_disp_grp)
shade_rows <- unlist(
  mapply(`:`,
         seq(3, n_rows, by = 4),
         seq(4, n_rows, by = 4))
)


sex_tbl <- sex_df_disp_grp %>%
kbl(
  caption = "Infection and severe infection rates by sex and Health Officer Region",
  align   = c("l", "l", "c", "c", "c", "c", "c"),
  format  = "html", 
  escape  = FALSE
) %>%
  
row_spec(
  0,
  bold = TRUE,
  background = "#0f172a",
  extra_css = "font-size: 16px!important;color:#ffffff;"
)%>%
  
column_spec(
  7,
  bold = TRUE,
  extra_css = "font-size: 14px!important;color:#0f172a;"
)%>%
  
row_spec(
  seq(2, n_rows, by = 2),
  extra_css = "border-bottom: 1.25px solid #0f172a;"
) %>%
  
row_spec(
    shade_rows,
    background = "#f1f0ea"
) %>%  
  
kable_styling(bootstrap_options = c("hover"), full_width = TRUE)

```